A Makefile for Developing Containerized \LaTeX\ Technical Documents

### 2020-12-24 ms.tex
reduce to 12 citations and remove footnotes

We propose a Makefile for developing containerized \LaTeX\ technical documents in POSIX-oriented operating systems.
By using portable technologies that withstood the test of time, we ensure the stability and portability of the proposed Makefile.
We write this paper\footnote{\url{https://github.com/pbizopoulos/a-makefile-for-developing-containerized-latex-technical-documents}}\footnote{\url{https://arxiv.org/abs/2005.12660}} using a template that consists of the Makefile\footnote{\url{https://github.com/pbizopoulos/a-makefile-for-developing-containerized-latex-technical-documents-template}} and choose Docker as the container technology and the Python programming language~\cite{van2007python} for generating the results.

@@

The proposed Makefile can be used for developing \textit{documents} that consist of \textit{results} generated from \textit{results code} executed within a container.
The three tools that are used by the Makefile are the Make program\footnote{\url{https://www.gnu.org/software/make/}} which is a build automation tool, \LaTeX~\cite{lamport1994latex} which is a typesetting system for publishing high quality research documents, and the container technology (such as Docker) which allows portable execution of \textit{code} between different hardware, operating systems and environments.

@@

Additionally, an \textit{author} can use the POSIX utility \textit{cmp} to compare \textit{documents} (generated in a different session, operating system or underlying hardware) byte by byte, and diffoscope\footnote{\url{https://diffoscope.org/}} for identifying sources of non-determinism as follows:
\begin{lstlisting}[language=bash, style=lststyle, caption={Test draft document version reproducibility. This can also be used as a test script when pushing or pull requesting to a remote repository.}]
# Generate the first draft document version.
make
# Backup the first resulting pdf.
cp tmp/ms.pdf tmp/ms-previous.pdf
# Update the modification timestamp of main.py.
touch main.py
# Generate the second draft document version.
make
# Compare draft document versions byte by byte.
cmp tmp/ms.pdf tmp/ms-previous.pdf
# Compare draft document versions
# in a human readable form.
diffoscope tmp/ms.pdf tmp/ms-previous.pdf
\end{lstlisting}

\section{Example use case of the Makefile}
This section provides a use case of the Makefile and also serves as a manual for developing \LaTeX\ \textit{documents} with Python as the \textit{results code} programming language.
We train, validate and test MobilenetV2 neural networks~\cite{sandler2018mobilenetv2} in PyTorch~\cite{paszke2019pytorch}, on MNIST~\cite{lecun2010mnist}, FashionMNIST~\cite{xiao2017fashion}, KMNIST~\cite{clanuwat2018deep}, CIFAR10~\cite{krizhevsky2009learning} and CIFAR100~\cite{krizhevsky2009learning}.
We use the default train, validation and test datasets and compare the use of ReLU, ReLU6\cite{dahl2013improving} and SiLU\cite{elfwing2018sigmoid} activation functions in the MobilenetV2 model.

@@

\begin{figure}[!t]
	\subfloat{\includegraphics[width=0.5\columnwidth]{tmp/MNIST-loss.pdf}}
	\subfloat{\includegraphics[width=0.5\columnwidth]{tmp/FashionMNIST-loss.pdf}}
	\caption{Comparison of the train and validation loss of MobilenetV2 with ReLU, ReLU6 and SiLU for MNIST and FashionMNIST.}
	\label{fig:loss}
\end{figure}

\begin{figure}[!t]
	\subfloat{\includegraphics[width=0.5\columnwidth]{tmp/CIFAR10-ReLU-kernels.pdf}}
	\subfloat{\includegraphics[width=0.5\columnwidth]{tmp/CIFAR100-ReLU-kernels.pdf}}
	\caption{The first 16 kernels of the first convolutional layer of MobilenetV2 with ReLU for CIFAR10 (left) and CIFAR100 (right).}
	\label{fig:kernels}
\end{figure}

@@

\section{Discussion}
The proposed Makefile is based on plain text file editing and can be used with any POSIX-oriented operating system that supports containers and Make.
POSIX and Make have existed for decades and have passed the test of time, while the container technology was standardized with the Open Container Initiative\footnote{\url{https://opencontainers.org/}} and there exist alternatives\footnote{\url{https://podman.io/}} that could be used as drop-in replacements for Docker.
Moreover the Makefile can be combined with any version control system, container registry provider and text editor thus preventing `app/vendor lock-in' situations.

>>>>

We propose a Makefile for developing containerized \LaTeX\ technical documents in POSIX-oriented operating systems.
By using portable technologies that withstood the test of time, we ensure the stability and portability of the proposed Makefile.
We write this paper using a template that consists of the Makefile and choose Docker as the container technology and the Python programming language for generating the results.

@@

The proposed Makefile can be used for developing \textit{documents} that consist of \textit{results} generated from \textit{results code} executed within a container.
The three tools that are used by the Makefile are the Make program which is a build automation tool, \LaTeX\ which is a typesetting system for publishing high quality research documents, and the container technology (such as Docker) which allows portable execution of \textit{code} between different hardware, operating systems and environments.

@@

Additionally, an \textit{author} can use the POSIX utility \textit{cmp} to compare \textit{documents} (generated in a different session, operating system or underlying hardware) byte by byte, and diffoscope for identifying sources of non-determinism as follows:
\begin{lstlisting}[language=bash, style=lststyle, caption={Test draft document version reproducibility. This can also be used as a test script when pushing or pull requesting to a remote repository.}]
# Generate the first draft document version.
make
# Backup the first resulting pdf.
cp tmp/ms.pdf tmp/ms-previous.pdf
# Update the modification timestamp of main.py.
touch main.py
# Generate the second draft document version.
make
# Compare draft document versions byte by byte.
cmp tmp/ms.pdf tmp/ms-previous.pdf
# Compare draft document versions
# in a human readable form.
diffoscope tmp/ms.pdf tmp/ms-previous.pdf
\end{lstlisting}

\section{Example use case of the Makefile}
This section provides a use case of the Makefile and also serves as a manual for developing \LaTeX\ \textit{documents} with Python as the \textit{results code} programming language.
We train, validate and test MobilenetV2 neural networks~\cite{sandler2018mobilenetv2} in PyTorch, on MNIST~\cite{lecun2010mnist}, FashionMNIST~\cite{xiao2017fashion}, KMNIST~\cite{clanuwat2018deep} and QMNIST~\cite{yadav2019cold}.
We use the default train, validation and test datasets and compare the use of ReLU, ReLU6\cite{dahl2013improving} and SiLU\cite{elfwing2018sigmoid} activation functions in the MobilenetV2 model.

@@

\begin{figure}[!t]
	\subfloat{\includegraphics[width=0.5\columnwidth]{tmp/MNIST-loss.pdf}}
	\subfloat{\includegraphics[width=0.5\columnwidth]{tmp/FashionMNIST-loss.pdf}}
	\\
	\subfloat{\includegraphics[width=0.5\columnwidth]{tmp/KMNIST-loss.pdf}}
	\subfloat{\includegraphics[width=0.5\columnwidth]{tmp/QMNIST-loss.pdf}}
	\caption{Comparison of the train and validation loss of MobilenetV2 with ReLU, ReLU6 and SiLU for all datasets.}
	\label{fig:loss}
\end{figure}

\begin{figure}[!t]
	\subfloat{\includegraphics[width=0.5\columnwidth]{tmp/MNIST-ReLU-kernels.pdf}}
	\subfloat{\includegraphics[width=0.5\columnwidth]{tmp/FashionMNIST-ReLU-kernels.pdf}}
	\\
	\subfloat{\includegraphics[width=0.5\columnwidth]{tmp/KMNIST-ReLU-kernels.pdf}}
	\subfloat{\includegraphics[width=0.5\columnwidth]{tmp/QMNIST-ReLU-kernels.pdf}}
	\caption{The first 25 kernels of the first convolutional layer of MobilenetV2 with ReLU for MNIST (upper-left), FashionMNIST (upper-right), KMNIST (lower-left) and QMNIST (lower-right).}
	\label{fig:kernels}
\end{figure}

@@

\section{Discussion}
The proposed Makefile is based on plain text file editing and can be used with any POSIX-oriented operating system that supports containers and Make.
POSIX and Make have existed for decades and have passed the test of time, while the container technology was standardized with the Open Container Initiative and there exist alternatives such as Podman that could be used as drop-in replacements for Docker.
Moreover the Makefile can be combined with any version control system, container registry provider and text editor thus preventing `app/vendor lock-in' situations.
